Parsing Postfix log files

### 2009-03-09 doc/logparser-architecture.tex
Removed the overly long cascaded parsing example.

    \item [Convert physical inputs to logical inputs] Each rule recognises
        a single input at a time: there is no facility for rules to consume
        more input or push unused input back onto the input stream,
        although actions may use cascaded parsing (\sectionref{cascaded
        parsing}) to effectively push input back onto the input stream.  A
        physical input (i.e.\ a single line read from the input stream) may
        contain multiple or partial logical inputs, and the framework must
        pre-process these to provide a logical input for the rules to
        recognise.  This simplifies the rules and actions considerably, at
        the cost of added complexity in the framework; during the design
        phase it was decided that it was easier to deal with the problem of
        parsing multiple or partial inputs once, rather than dealing with
        it in every rule and action.  This is easy in the case of Postfix
        log files because there is a one-to-one mapping between physical
        and logical inputs; mapping between physical and logical inputs may
        be more difficult for other types of input.  Some input types may
        require pre-processing equivalent to parsing the physical inputs;
        in those cases the framework should behave similarly to other
        architectures: present the entire input stream at once, and discard
        the portion recognised by the successful rule.

@@

    \item [Invoke actions]  Once a rule has been found that recognises the
        current input, the specified action must be invoked.  The framework
        marshals the data extracted by the rule, invokes the action, and
        repeats the parsing process if cascaded parsing
        (\sectionref{cascaded parsing}) is used.

@@

\subsection{Cascaded Parsing}

\label{cascaded parsing}

Actions may return a modified input line to the framework that will be
parsed as if read from the input stream, allowing for a simplified version
of cascaded parsing~\cite{cascaded-parsing}.  This powerful facility allows
several rules and actions to parse a single input, potentially simplifying
both rules and actions.  A simple example is to have one rule and action
removing comments from input lines, so that other rules and actions do not
have to handle comments at all; obviously if comment characters can be
escaped or embedded in quoted strings the implementation needs to be more
complex, but for some inputs this can greatly simplify subsequent parsing.
A more complex example is evaluating simple arithmetic expressions, where
sub-expressions enclosed in parentheses must be evaluated first; cascaded
parsing can be used to parse and evaluate sub-expressions, substituting the
result into the original expression.  Actions do not need to be specially
registered with the framework or be declared in a particular way to use
cascaded parsing: actions that do not use cascaded parsing will return
nothing, those that do will simply return a string to be parsed.

XXX MAYBE MOVE THIS EXAMPLE TO AN APPENDIX\@?  DO I EVEN NEED IT\@?  ASK
CARL ABOUT IT\@.

XXX GET GREG TO PROOFREAD THIS SECTION\@.

A more complicated example is determining the atomic weight of a compound
given its chemical formula.  A chemical formula is a sequence of chemical
element abbreviations, each optionally followed by a number indicating how
many of that element are present in the formula (if omitted the number is
assumed to be one).  Some parts of a formula may be repeated: they are
enclosed in parentheses, and the closing parenthesis is followed by a
number indicating how many are present in the formula.  A monomer is a
simple compound, i.e.\ it does not have any repeated portions; compounds
with repeated portions are called polymers.  E.g.\ glucose, a monomer, has
the chemical formula $C_{6}H_{12}O_{6}$, and hydrocarbon, a polymer, is
$CH_{3}{(CH_{2})}_{50}CH_{3}$.  The atomic weight of an individual element
is given in periodic tables, but one element in particular will be
important in the later explanation: Hydrogen, abbreviated H, has an atomic
weight of 1.  The atomic weight of a given compound is the sum of the
atomic weights of each of its elements multiplied by the number of times
they occur; this is simple to calculate for a monomer, but it is harder for
a polymer because the atomic weight of the repeated monomer must be
multiplied by the number of times the monomer is present in the compound
(e.g.\ for hydrocarbon the atomic weight of $CH_{2}$ must be multiplied by
$50$).

Calculating the atomic weight of a monomer requires a single rule and
action.  The action separates each element in the monomer, looks up its
atomic weight, multiplies it by the number of times it is present (one if
unspecified), and adds up the results.  The rule's regex is
\verb!^([A-Z]([a-z])?\d*)+$!, which does not guarantee that the chemical
element abbreviations in the formula are correct; it could be modified to
check that by replacing \verb![A-Z]([a-z])?! with all the valid
abbreviations, slightly simplifying the action at the cost of complicating
the regex, which could be automatically generated.

Parsing a polymer is more difficult because of the repeated monomers: each
monomer's atomic weight needs to be calculated separately, multiplied by
the number of times it is present in the compound, and that figure added to
the sum.  Without cascaded parsing the action would need to maintain state
while recursively parsing the monomers embedded in the polymer.  Using
cascaded parsing, three rules and actions can be used instead as shown in
\tableref{Regexes and Actions to calculate the atomic weight of a chemical
compound}.

% These figures come from experimentation; set the length to 0pt, see how
% big a hbox latex complains about, and then set the length.  Stupid latex
% cannot do \settowidth{\verb}.  Poo.
\newlength{\chemicalactionwidth}
\setlength{\chemicalactionwidth}{\textwidth}
% This is the width of the number column, roughly.
\addtolength{\chemicalactionwidth}{-30pt}
% The width of the regex column.
\addtolength{\chemicalactionwidth}{-130pt}
% This is the overhead of space around columns, on either side, etc.
\addtolength{\chemicalactionwidth}{-24pt}

\begin{table}[ht]
    \caption{Regexes and Actions to calculate the atomic weight of a
    chemical compound}
    \empty{}\label{Regexes and Actions to calculate the atomic weight of a
    chemical compound}
    \begin{tabular}{llp{\chemicalactionwidth}}

        \tabletopline{}%
        No.   & Regex                       & Action \\
        \tablemiddleline{}%
        1     & \verb'((?!H\d+([()]|$))'    & Parse a monomer as described above,  \\
              & \verb' ([A-Z]([a-z])?\d*)+)'& returning \verb!Hnnn!, where \verb!nnn! is
                                              the atomic weight of the monomer.  The first
                                              part of the regex prevents it from parsing
                                              the return value of this action
                                              (\verb!Hnnn!), ensuring the parsing process
                                              does not get stuck in a loop.  \\
        2     & \verb!\(H(\d+)\)(\d+)!      & Parse a repeated monomer that has been
                                              rewritten by the first rule and action,
                                              returning \verb!Hnnn!, where \verb!nnn! is
                                              the product of the first and second numbers
                                              captured by the regex.  \\
        3     & \verb!^H(\d+)+$!            & Repeated applications of rules and
                                              actions one and two will have rewritten the
                                              input to \verb!Hnnn!, where \verb!nnn! is
                                              the atomic weight of the compound.  This
                                              action should do whatever is required with
                                              the calculated atomic weight.  \\
        \tablebottomline{}%

    \end{tabular}
\end{table}

Applying the rules and actions above to the hydrocarbon formula results in
this sequence:

\begin{tabular}{lrl}

    1 & Start with              & $CH_{3}{(CH_{2})}_{50}CH_{3}$ \\
    2 & Apply rule and action 1 & $H_{15}{(CH_{2})}_{50}CH_{3}$ \\
    3 & Apply rule and action 1 & $H_{15}{(H_{14})}_{50}CH_{3}$ \\
    3 & Apply rule and action 1 & $H_{15}{(H_{14})}_{50}H_{15}$ \\
    5 & Apply rule and action 2 & $H_{15}H_{700}H_{15}$ \\
    6 & Apply rule and action 1 & $H_{730}$ \\
    7 & Apply rule and action 3 & 730 \\

@@

\newpage{} % XXX
\subsection{Attaching Conditions To Rules}

>>>>

    \item [Convert physical inputs to logical inputs] Each rule recognises
        a single input at a time: there is no facility for rules to consume
        more input or push unused input back onto the input stream,
        although actions may use cascaded parsing (explained in
        \sectionref{actions in architecture}) to effectively push input
        back onto the input stream.  A physical input (i.e.\ a single line
        read from the input stream) may contain multiple or partial logical
        inputs, and the framework must pre-process these to provide a
        logical input for the rules to recognise.  This simplifies the
        rules and actions considerably, at the cost of added complexity in
        the framework; during the design phase it was decided that it was
        easier to deal with the problem of parsing multiple or partial
        inputs once, rather than dealing with it in every rule and action.
        This is easy in the case of Postfix log files because there is a
        one-to-one mapping between physical and logical inputs; mapping
        between physical and logical inputs may be more difficult for other
        types of input.  Some input types may require pre-processing
        equivalent to parsing the physical inputs; in those cases the
        framework should behave similarly to other architectures: present
        the entire input stream at once, and discard the portion recognised
        by the successful rule.

@@

    \item [Invoke actions]  Once a rule has been found that recognises the
        current input, the specified action must be invoked.  The framework
        marshals the data extracted by the rule, invokes the action, and
        repeats the parsing process if cascaded parsing (see
        \sectionref{actions in architecture}) is used.

@@

Actions may return a modified input line to the framework that will be
parsed as if read from the input stream, allowing for a simplified version
of cascaded parsing~\cite{cascaded-parsing}.  This powerful facility allows
several rules and actions to parse a single input, potentially simplifying
both rules and actions.  A simple example is to have one rule and action
removing comments from input lines, so that other rules and actions do not
have to handle comments at all; obviously if comment characters can be
escaped or embedded in quoted strings the implementation must be careful
not to remove those.  For some inputs this type of pre-processing can
greatly simplify parsing, echoing the simplification provided by the
framework presenting rules and actions with logical inputs rather than
physical inputs.  A more complex example use of cascaded parsing is
evaluating simple arithmetic expressions, where sub-expressions enclosed in
parentheses must be evaluated first; cascaded parsing can be used to parse
and evaluate the sub-expressions, substituting the result into the original
expression for subsequent re-evaluation.  Actions do not need to be
specially registered with the framework or be declared in a particular way
to use cascaded parsing: actions that do not use cascaded parsing will
return nothing, those that do will simply return a string to be parsed.

@@

\subsection{Attaching Conditions To Rules}
